Preserving Reciprocal Consistency in Distributed Graph Databases

### 2020-04-05 paper/main.tex
fix up

\usepackage{subcaption}
\usepackage{tikz}
\usepackage{fancyvrb}
\usepackage{xcolor}
\usepackage{breqn}

@@

Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}. Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}. Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities. In addition, properties can be stored on both vertices and edges. In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects. This allows for bi-directional traversal and improved query performance \cite{Robinson2015}. An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (Details in Section \ref{sec:recipr-cons}).

In the wild graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases. A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions. Then, a BASE database \cite{Pritchett2008} is used for storage and adapted with a graph processing layer (\cite{janusgraph}, \cite{TitanDB}). A caveat with this approach is BASE databases often eschew transactional guarantees in order to achieve higher performance, specifically there is often no concurrency control mechanism for operations that spans partitions.

Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted how reciprocal consistency of distributed edges can be violated in such a design, introducing corruption into the database. Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates. Preserving reciprocal consistency of distributed edges requires some degree of coordination between partitions in the presence of concurrent modifications. This paper proposes and analyses a lightweight protocol that ensures reciprocal consistency of distributed edges in operational contexts where traditional concurrency control mechanisms have been done away with for sake of performance.

@@

A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both data values and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}. In this representation an edge has \emph{reciprocal} entries in the adjacency lists of the vertices the edge connects. A query reading either the source or destination vertex should be able to reify the edge correctly, returning consistent results. When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
  \caption{Database records representing vertices, containing data values and adjacency lists \cite{janusgraph}.}
  \label{adj-list}
\end{figure}

Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit. It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).

@@

Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}'. Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction.

\section{Distributed Graph Databases}

@@

However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database. BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions. This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}. Earlier work investigated how the lack of concurrency control across partitions in BASE distributed graph databases can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}). This process is explained in Section \ref{sec:db-corruption}.

\begin{figure*}[ht]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}
      \draw [thick] (1.3,3) -- (1.7,3);
      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
      \draw [thick] (3.3,3) -- (3.7,3);
      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
    \end{tikzpicture}
    \caption{}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}
      \draw [thick] (5.3,3) -- (5.7,3);
      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
      \draw [thick] (7.3,3) -- (7.7,3);
      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
    \end{tikzpicture}
    \caption{}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}
      \draw [thick] (9.3,3) -- (9.7,3);
      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
      \draw [thick] (11.3,3) -- (11.7,3);
      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
    \end{tikzpicture}
    \caption{}
  \end{subfigure}%
  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$. In (a) $T_y$ begins writing to the distributed edge before $T_x$, in (b) the converse is true, else they are equivalent. In (c) both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
  \label{conf-scen}
\end{figure*}

\section{Corruption in BASE  Distributed Graph Databases}
\label{sec:db-corruption}

When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively. Without sufficient concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated, Figure \ref{conf-scen}. For such edges there exists a correct and a incorrect entry. Note, the order in which the two writes take place is not constrained. For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}. A graph with half-corrupted edges has suffered \emph{structural corruption}.

@@

\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
\textcolor{grey}{// Tz}
\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
\textcolor{blue}{AND} b.\textcolor{red}{name} = 'unknown'
\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
\end{Verbatim}

    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);

  \end{tikzpicture}
  \caption{A half-corrupted edge.}
  \label{half-corrupted}
\end{figure}

\section{\emph{Delta} Protocol}

A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}. Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.

@@

This approach is problematic in a distributed graph database as a subset of distributed edges are traversed and modified with a high frequency e.g. critical sections of motorway in a road network, leading to high contention. In addition, graph transactions tend to be longer-lived than transactions in relational databases. Taking long duration locks on these frequently accessed edges significantly limits concurrency and reduces throughput. With these concerns in mind we developed the \emph{Delta} protocol, which aimed at preventing distributed edges becoming half-corrupted and hence quashing the seed of corruption. We also wanted to keep performance at levels that make using a BASE distributed graph database an attractive choice.\footnote{The $\Delta$ protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}

The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end. It is then assumed that the network delay between two servers can be predicted to be some $\Delta$. All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed. From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time. Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$. An aborted transaction, aborts any and every previous tentative write that it may have successfully completed. This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.

This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency. The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased. However, if $\Delta$ is sufficiently large (larger than the duration of the transaction) the protocol produces outcomes equivalent to taking long-duration write locks on distributed edges with \texttt{NO_WAIT} deadlock detection.

% Consider two transactions $T_1$ and $T_2$ write two distributed edges $X$ and $Y$:

%  $\mathbf{T_1}: W_s(X), W_d(X), W_d(Y), W_s(Y)$ \\
%  $\mathbf{T_2}: W_s(Y), W_d(Y), W_s(X), W_d(X)$

%  Here $T_1$ and $T_2$ can interleave in a manner where for $X, T_1 \rightarrow T_2 $ and for $Y, T_2 \rightarrow T_1$.
\begin{figure}[H]
  \centering
  \begin{tikzpicture}[node distance=2cm]
    % s1
    \draw [thick] (1.3,3) -- (1.7,3);
    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
    % s2
    \draw [thick] (3.3,3) -- (3.7,3);
    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};

    \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
    \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);

    \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};

  \end{tikzpicture}
  \caption{An example of the \emph{Delta} protocol preserving reciprocal consistency.}
  \label{delta-abort}
\end{figure}

Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption. Figure \ref{corruption-again} displays an example when the \emph{Delta} protocol fails to maintain reciprocal consistency. The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.

\begin{figure}[H]
  \centering
  \begin{tikzpicture}[node distance=2cm]
    \draw [thick] (1.3,3) -- (1.7,3);
    \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
    \draw [thick] (3.3,3) -- (3.7,3);
    \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
        \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
    \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};

    \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
    \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
  \end{tikzpicture}
  \caption{An example of the \emph{Delta} protocol failing to preserve reciprocal consistency.}
  \label{corruption-again}
\end{figure}

In summary, the $\Delta$ protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded operational corruption can still occur. Two questions naturally arise regards the performance of the protocol. For different values of $\Delta$:
\begin{itemize}
\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
\item How many transactions are aborted as a result of the protocol?
\end{itemize}

@@

The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second. To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write. To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$. For each type, a fraction $f$ are distributed edges and the remainder are local edges. The network delay between servers was assumed to be exponentially distributed with rate $\rho$. At any moment in time an edge can be in one of four states:
\begin{enumerate}
\item Local and clean.
\item Distributed and clean.
\item Half-corrupted distributed edge arising from a dirty write.
\item Semantically corrupted.
\end{enumerate}
The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.

>>>>

\usepackage{xspace}
\usepackage[labelformat=simple]{subcaption}
\renewcommand\thesubfigure{(\alph{subfigure})}
\usepackage{tikz}
\usepackage{fancyvrb}
\usepackage{xcolor}
\usepackage{breqn}

\newcommand{\tDelta}{\textsf{Delta}\xspace}

@@

Recent years have seen a proliferation in the use of graph processing technologies \cite{Besta2019}. Application areas are wide reaching from healthcare, to social networks and fraud detection \cite{Eifrem2016}. Graph databases model data as a \textit{property graph} \cite{Robinson2015}, vertices represent entities and edges represent the relationships between entities. In addition, properties can be stored on both vertices and edges. In the storage layer, edges are represented by two reciprocal pointers, one stored with each vertex the edge connects. This allows for bi-directional traversal and improved query performance \cite{Robinson2015}. An edge is said to be \emph{reciprocally consistent}, if its two end pointers are mutually reciprocal of each other (details in Section \ref{sec:recipr-cons}).

In practice, graphs can be extremely large, sometimes in the magnitude of 100 billion edges \cite{Sahu2017}, exceeding the storage capacity of a single-node graph database and motivating the need for distributed graph databases. A common distributed graph database design pattern is to first partition graph data over several machines in a cluster; resulting in a number of \emph{distributed edges}, an edge's reciprocal pointers reside in different partitions. Recent work \cite{Ezhilchelvan2018} and \cite{Webber2019} highlighted that violations of reciprocal consistency in distributed edges introduce corruption into the database. Moreover, due to the \emph{Scale-Free} \cite{ScaleFree} property exhibited by many real world graphs, this corruption can propagate through the database at alarmingly rates.

When, for example, a BASE database \cite{Pritchett2008} is adapted with a graph processing layer, then violations of reciprocal consistency will occur if that adaptation provides no concurrency control for operations that span partitions in order to offer higher performance. This paper proposes a simple concurrency control protocol, called \tDelta protocol, that does not impede performance adversely. That is because the protocol is exclusively designed for one purpose only: reciprocal consistency in distributed edges.  Its design leverages the fact that the updating of end pointers of a given distributed edge immediately follow each other and the small interval between them is the window for possible conflicts between concurrent updates. The protocol ensures that any two updates on any given end pointer of any given edge are at least $\Delta$ time apart by aborting any write attempt that is within . It ensures reciprocal consistency when $\Delta$ is chosen to be larger than the length of the conflict window.

The paper is organized as follows: Section \ref{sec:recipr-cons} presents reciprocal consistency, Sections \ref{sec:distr-graph-datab}-\ref{sec:db-corruption} discuss the architecture of distributed graph databases and describes how corruption occurs. Section \ref{sec:tdelta-protocol} outlines the \tDelta protocol. Sections \ref{sec:modeling}-\ref{sec:evaluation} presents the model used to evaluate protocol performance.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\linewidth]{./images/janusgraph-adj-list}
  \caption{Database records representing vertices, containing properties and a sequence of edges that comprise the adjacency list of a vertex \cite{janusgraph}.}
  \label{adj-list}
\end{figure}

@@

A common approach to storing graphs (arising from JanusGraph \cite{janusgraph} and TitanDB \cite{TitanDB}) is for database records to represent vertices containing both properties and an adjacency list containing edge \emph{pointers} to other vertices, Figure \ref{adj-list}. Consider, for example, the statement that Tolkien \textit{wrote} The Hobbit. It is expressed using vertices \emph{a} and \emph{b}, for Tolkien and The Hobbit respectively, and an edge \textit{wrote} running from \emph{a} (source) to \emph{b} (destination).

@@

Adjacency lists of both \emph{a} and \emph{b} record information about the edge and this information is mutually reciprocal (or inverse) of each other: \emph{a}'s list will indicate `\emph{a} \emph{wrote} \emph{b}' while \emph{b}'s will have `\emph{b} \emph{written} by \emph{a}', Figure \ref{reciprocally-consistent-edge}. Thus, a query `list all titles by the author who wrote The Hobbit' can be answered starting at (destination vertex) \emph{b} and then traversing to (source vertex) \emph{a}, even though the edge is ``directed'' from \emph{a} to \emph{b} at model level abstraction. When the adjacency list entries for a given edge are mutually compatible like this, that edge is said to be \emph{reciprocally consistent}, a form of referential integrity. A query can read either the source or destination vertex and is able to reify the edge correctly, returning consistent results.

    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} (4.2,0);
  \end{tikzpicture}
  \caption{A reciprocally consistent edge.}
  \label{reciprocally-consistent-edge}
\end{figure}

\section{Distributed Graph Databases}
\label{sec:distr-graph-datab}

@@

However, the problem with this design is the (lack of) transactional semantics is inherited from the underlying database. BASE databases seldom provide guarantees for multi-operation, multi-object transactions that span partitions. This lack of concurrency control across partitions makes it possible for concurrent updates to interleave in a manner that violates reciprocal consistency of distributed edges\footnote{When operating without concurrency control it is equally possible that local edges become reciprocal consistent, however primitives provided by BASE databases are typically sufficient to ensure reciprocal consistency for local edges.}. Earlier work investigated how the lack of concurrency control across partitions can undermine reciprocal consistency of distributed edges, causing irreversible corruption that spreads at alarmingly rates (\cite{Ezhilchelvan2018}, \cite{Webber2019}). This process is explained in Section \ref{sec:db-corruption}.

\begin{figure*}[ht]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}
      \draw [thick] (1.3,3) -- (1.7,3);
      \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
      \draw [thick] (3.3,3) -- (3.7,3);
      \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
      \draw [thick,<-,>=stealth] (1.5,1.5) -- (3.5,2.5)  node[right] {$T_y$};
      \draw [thick,->,>=stealth] (1.5,2.2) node[left] {$T_x$} -- (3.5,1);
    \end{tikzpicture}
    \caption{}
    \label{fig:tx-ty}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}
      \draw [thick] (5.3,3) -- (5.7,3);
      \draw [thick,<-,>=stealth] (5.5,0) -- (5.5,3) node[anchor=south] {$S_i$};
      \draw [thick] (7.3,3) -- (7.7,3);
      \draw [thick,<-,>=stealth] (7.5,0) -- (7.5,3) node[anchor=south] {$S_j$};
      \draw [thick,<-,>=stealth] (5.5,1) -- (7.5,2)  node[right] {$T_y$};
      \draw [thick,->,>=stealth] (5.5,2.7) node[left] {$T_x$} -- (7.5,1.2);
    \end{tikzpicture}
    \caption{}
\label{fig:ty-tx}
  \end{subfigure}%
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \begin{tikzpicture}
      \draw [thick] (9.3,3) -- (9.7,3);
      \draw [thick,<-,>=stealth] (9.5,0) -- (9.5,3) node[anchor=south] {$S_i$};
      \draw [thick] (11.3,3) -- (11.7,3);
      \draw [thick,<-,>=stealth] (11.5,0) -- (11.5,3) node[anchor=south] {$S_j$};
      \draw [thick,->,>=stealth] (9.5,2.5)  node[left] {$T_x$} -- (11.5,1) ;
      \draw [thick,->,>=stealth] (9.5,2) node[left] {$T_y$} -- (11.5,1.7);
    \end{tikzpicture}
    \caption{}
    \label{fig:same-side}
  \end{subfigure}%
  \caption{Possible interleavings of concurrent transaction's writes to a distributed edge spanning servers $S_i$ and $S_j$ by transactions $T_x$ and $T_y$. In \subref{fig:tx-ty} $T_y$ begins writing to the distributed edge before $T_x$, in \subref{fig:ty-tx} the converse is true, else they are equivalent. In \subref{fig:same-side} both transactions begin writing at the same server but overlap in the network and arrive out-of-order.}
  \label{conf-scen}
\end{figure*}

\begin{figure}[H]
  \centering
  \begin{tikzpicture}[node distance=2cm]
    \node (v1) [vertex2,xshift=0cm,yshift=0cm] {\small{\texttt{a:\textcolor{green}{Person}}}};

    \draw [thick,->,>=stealth] (0.8,0)  -- node [midway,above] {:\textcolor{green}{\small{\texttt{WROTE}}}} node [midway,below] {\small{\texttt{\textcolor{red}{year} = 1937}}} (2.4,0);
  \end{tikzpicture}
  \caption{A half-corrupted edge.}
  \label{half-corrupted}
\end{figure}

\section{Corruption in the Absence of Concurrency Control}
\label{sec:db-corruption}

When a given transaction writes a distributed edge it must infact perform two writes, writing reciprocal entries in the adjacency lists at the source and the destination vertices, which say reside on servers $S_i$ and $S_j$ respectively. Without concurrency control, transaction's writes to a distributed edge can interleave producing a distributed edge in a \emph{half-corrupted} state - reciprocal consistency has been violated. Possible interleavings of concurrent writes to a distributed edge are given in Figure \ref{conf-scen}. A half-corrupted edge is an example of a \emph{dirty write} (ANSI \emph{P0} \cite{Berenson1995}, Adya \emph{G0} \cite{Adya2000}), which is proscribed by the weakest ANSI isolation level \textbf{Read Uncommitted}. Under Read Uncommitted the database ensures a total order on transactions, consistently ordering writes from concurrent transactions, which would prevent all interleavings in Figure \ref{conf-scen}.

For such half-corrupted edges there exists a correct and a incorrect entry. Note, the order in which the two writes take place is not constrained. For example, when updating an edge between \emph{a} and \emph{b} it is equally likely to update \emph{a} then \emph{b} as it is to update \emph{b} then \emph{a}. A graph with half-corrupted edges has suffered \emph{structural corruption}.

@@

\begin{Verbatim}[commandchars=\\\{\},fontsize=\small,xleftmargin=.2in]
\textcolor{grey}{// Tz}
\textcolor{blue}{MATCH} (b:\textcolor{green}{Book}),(u:Person)
\textcolor{blue}{WHERE NOT} (:\textcolor{green}{Person})-[:\textcolor{green}{WROTE}]->(b:\textcolor{green}{Book})
\textcolor{blue}{AND} u.\textcolor{red}{name} = 'unknown'
\textcolor{blue}{CREATE} (u)-[:\textcolor{green}{WROTE}]->(b)
\end{Verbatim}

\section{\tDelta Protocol}
\label{sec:tdelta-protocol}

@@

Our \tDelta protocol employs principles behind all these well-tested strategies but has two crucial differences:
\begin{itemize}
\item No locks are used.
\item A write operation need not await until the preceding write commits but can proceed if at least $\Delta$ duration (measured in local clock) has elapsed.
\end{itemize}

These differences lead to several advantages in the context of graph databases. Firstly, a subset of edges are traversed and modified with a high frequency e.g. critical sections of motorway in a road network, leading to high contention. Secondly, graph transactions tend to be longer-lived than transactions in relational databases. Waiting for earlier writes on edges by long running transactions to commit, these frequently accessed edges significantly limit concurrency and reduce throughput. With these concerns in mind we developed the \tDelta protocol, which aimed at preventing edges becoming half-corrupted and hence quashing the seed of corruption whilst keeping performance at an acceptable level.\footnote{The \tDelta protocol is solely a concurrency control mechanism for distributed edges, guarantees about vertices, local edges are beyond the scope of this paper.}

\subsection{Protocol Description}
\label{sec:protocol-description}

The \tDelta Protocol has five rules:
\begin{enumerate}
\item A transaction's update on an end pointer of an edge is initially tentative which would become permanent only if that transaction is permitted to commit.
\item A tentative write is possible if the end pointer is either in a permanent state or the immediately preceding tentative update was done at least $\Delta$ time before (where time is measured as per local clock).
\item If a transaction successfully performs all its tentative updates, then it is permitted to commit; otherwise, it must abort.
\item A transaction commits, when all its tentative updates are made permanent, e.g., using an atomic commitment protocol.
\item Tentative updates of an aborting transactions are ignored. An ignored tentative update can make a new transaction abort for up to $\Delta$ time after it was created; it is harmless thereafter and can be garbage collected at any time.
\end{enumerate}

\subsection{Correctness Reasoning}
\label{sec:corr-reas}

Let us define $\delta$ as the bound estimate on the interval that may elapse between a transaction updating one end of a distributed edge at one server and updating the other end of the same edge at another server.

Let $\Delta$ be chosen such that $\Delta > \delta$. Now consider the interleaving in Figure \ref{fig:tx-ty} and let $t_x$ be the (global) time when $T_x$ starts at server $S_i$; similarly $t_y$ be the time $T_y$ starts at server $S_j$. Note $t_y < t_x$, i.e. $(t_x - t_y) > 0$.

Say, $T_y$ reaches $S_i$ at time $t_y + d$, where $d$ is the actual time elapsed between completing a tentative write at one end and starting at another end, let us assume that $d \leq \delta$. When $T_y$ arrives at $S_i$ it will find a tentative write done at time $t_x$. In this case, $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$; so, $T_y$ will abort, preventing writes from interleaving and half-corrupting the edge. A similar arguments can be made for the scenario in with Figure \ref{fig:ty-tx}. For the interleaving in Figure \ref{fig:same-side}, if $t_y - t_x > \Delta$, then $T_y$ cannot overtake $T_x$ at $S_j$.

Assume that $d > \Delta$, i.e., the estimate $\delta$ does not hold at this moment. In Figure \ref{fig:tx-ty} interfering updates are avoided only if $t_y + d - t_x = d - (t_x-t_y) < d \leq \delta < \Delta$ otherwise, an edge becomes half-corrupted. This holds for the interleaving in \ref{fig:ty-tx}. Thus, in the extreme case $t_x = t_y$, reciprocal consistency is not guaranteed if $d > \Delta$ when the former exceeds its upper bound estimate $\delta$. For Figure \ref{fig:same-side}, if $d$ for $T_x$ is larger than $\Delta$, reciprocal consistency can be violated.

This protocol is strictly weaker than Read Uncommitted isolation, ensuring only writes to distributed edge are totally ordered, preserving reciprocal consistency. The benefit of this approach is reduced contention on frequently accessed distributed edges, as the time a transaction has exclusive access to a given distributed edge is decreased.

In summary, the \tDelta protocol attempts to prevent the occurrence of half-corrupted distributed edges by aborting transactions but if $\Delta$ is exceeded edge can become half-corrupted and process of corruption described in Section \ref{sec:db-corruption} can still occur. Two questions naturally arise regards the performance of the protocol. For different values of $\Delta$:
\begin{itemize}
\item Given distributed edges can still become half-corrupted if $\Delta$ is exceeded, by how much time does the protocol increase the time until operational corruption?
\item How many transactions are aborted as a result of the protocol?
\end{itemize}

%The two paragraphs (The protocol leverages... edge is lost) can be commented out.

%Figures 5 and 6 can be removed.

% The protocol leverages the fact that a transaction that writes at one end of a distributed edge, must then immediately write the other end. It is then assumed that the network delay between two servers can be predicted to be some $\Delta$. All writes are temporary and assumed to be made permanent by a later atomic commitment protocol once the transaction has completed. From this, a distributed edge write is allowed to proceed provided there is no other temporary write preceding it within $\Delta$; measured as per the local clock time. Figure \ref{delta-abort} shows an instance of the protocol preventing the interleaving in Figure \ref{conf-scen} (c), $T_x$ overtakes $T_y$ but is aborted when arriving at $S_j$ as $T_y$ has already wrote tentatively within $\Delta$. An aborted transaction, aborts any and every previous tentative write that it may have successfully completed. This protocol avoids all conflict scenarios in Figure \ref{conf-scen} provided $\Delta$ is not exceeded.

% \begin{figure}[H]
%   \centering
%   \begin{tikzpicture}[node distance=2cm]
%     % s1
%     \draw [thick] (1.3,3) -- (1.7,3);
%     \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
%     % s2
%     \draw [thick] (3.3,3) -- (3.7,3);
%     \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};

%     \draw [thick,->,>=stealth] (1.5,2.5) node[left] {$T_x$} -- (3.5,0.5);
%     \draw [thick,->,>=stealth] (1.5,2) node[left] {$T_y$} -- (3.5,1.8);

%     \draw [thick,<->,>=stealth] (3.7,0.5) -- (3.7,1.8) node[midway,right] {$< \Delta$};

%   \end{tikzpicture}
%   \caption{An example of the \tDelta protocol preserving reciprocal consistency.}
%   \label{delta-abort}
% \end{figure}

% Correctly choosing $\Delta$ is paramount, as if it is exceeded all three conflict scenarios can still occur, resulting in half-corrupted edges and the spread of semantic corruption. Figure \ref{corruption-again} displays an example when the \tDelta protocol fails to maintain reciprocal consistency. The network delay between $S_i$ and $S_j$ is sufficiently large the total ordering of writes to the distributed edge is lost.

% \begin{figure}[H]
%   \centering
%   \begin{tikzpicture}[node distance=2cm]
%     \draw [thick] (1.3,3) -- (1.7,3);
%     \draw [thick,<-,>=stealth] (1.5,0) -- (1.5,3) node[anchor=south] {$S_i$};
%     \draw [thick] (3.3,3) -- (3.7,3);
%     \draw [thick,<-,>=stealth] (3.5,0) -- (3.5,3) node[anchor=south] {$S_j$};
%         \draw [thick,->,>=stealth] (1.5,2.7) node[left] {$T_x$} -- (3.5,0.8);
%     \draw [thick,<-,>=stealth] (1.5,1) -- (3.5,2.2)  node[right] {$T_y$};

%     \draw [thick,<->,>=stealth] (4.2,1) -- (4.2,2.2) node[midway,left] {$> \Delta$};
%     \draw [thick,<->,>=stealth] (0.8,1) -- (0.8,2.5) node[midway,right] {$> \Delta$};
%   \end{tikzpicture}
%   \caption{An example of the \tDelta protocol failing to preserve reciprocal consistency.}
%   \label{corruption-again}
% \end{figure}

@@

The system processes transactions that arrive in a Poisson stream with rate $\lambda$ per second. To simplify the model, each transaction contains a random number of read operations, $K$, followed by a single write. To model a scale-free graph, edges in the database are divided into $T$ types, popular edges types have higher access probabilities but are a smaller proportion of the total number of edges, $N$. For each type, a fraction $f$ are distributed edges and the remainder are local edges. The network delay between servers was assumed to be exponentially distributed with rate $\rho$. At any moment in time an edge can be in one of four states:
\begin{enumerate}
\item Local and clean.
\item Distributed and clean.
\item Half-corrupted distributed edge arising from a dirty write.
\item Semantically corrupted.
\end{enumerate}
% The valid state transitions are given in Figure \ref{state-transitions}. Note, only distributed edges can be in state 2, but any edge, including local ones, can be in state 3.

% \begin{figure}[ht]
%   \centering
%     \begin{tikzpicture}[node distance=2cm]
%       \node (n0) [circle,draw,radius=0.3, align=center,green] {0};
%       \node (n1) [circle,draw,radius=0.3, align=center, right of=n0,green] {1};
%       \node (n2) [circle,draw,radius=0.3, align=center, right of=n1,orange] {2};
%       \node (n3) [circle,draw,radius=0.3, align=center, right of=n2,red] {3};
%       \draw [thick,<-,>=stealth,green] (n1) to[out=-45,in=-135]  node[below] {$a_{2,1}$} (n2);
%       \draw [thick,->,>=stealth,orange] (n1)  to[out=0,in=-180] node[below] {$a_{1,2}$} (n2);
%       \draw  [thick,->,>=stealth,red] (n2) to[out=-45,in=-135]  node[below] {$a_{2,3}$} (n3);
%       \draw  [thick,->,>=stealth,red]  (n0) to[out=70,in=110] node[above] {$a_{0,3}$} (n3);
%       \draw  [thick,->,>=stealth,red]  (n1) to[out=45,in=135] node[above] {$a_{1,3}$} (n3);
%     \end{tikzpicture}
%     \caption{Edge transitions between clean, half-corrupted and semantically corrupt states.}
%     \label{state-transitions}
% \end{figure}
